We are not the 99 percent: quantifying asphericity in the distribution of Local Group satellites

### 2017-11-22 paper/paper.tex
new plot
abstract

\begin{abstract}
Here we quantify the asphericity in the spatial distribution of the
brightest satellites around the Milky Way (MW) and M31, the dominant
galaxies in the Local Group (LG).
We use the Illustris-1 and ELVIS simulations to quantify asphericity
in the Lambda Cold Dark Matter paradigm.  
We build explicit probability distributions of
finding an aspherical satellite distribution, allowing for an easy
reproducibility of the results presented here and a standarized method
to compare probability distributions derived from different
simulations. 
We estimate that only $0.05\%$ to $0.25\%$ of the pairs with isolation
and dynamical characteristics similar to the LG are expected to have
satellite distributions with the same degree of deviation from
sphericity.
This makes the LG a $3\sigma$ outlier, with all the weight lying on
the extremly planar MW satellite distribution.

@@

\begin{figure}
\centering
\includegraphics[width=0.47\textwidth]{expected_numbers.pdf}
\caption{Probability distribution for the expected number of pairs
  showing the same degree of atipicality as the Local Group if drawn
  from a sample of 2000 isolated pairs. 
  The distributions correspond to results derived from Illustris-1
  and ELVIS data.
  On average between $0.05\%$ and $0.25\%$ of the isolated pairs should present
  satellite distributions as atypical as the Local Group.
\label{fig:expected_number}}
\end{figure}

>>>>

\begin{abstract}
Here we quantify the asphericity in the spatial distribution of the
brightest satellites around the Milky Way (MW) and M31, the dominant
galaxies in the Local Group (LG).
We build explicit probability distributions of finding an aspherical
satellite distribution. 
This allows us to estimate the atipicallity of the satellite
distributions in the LG even when the underlying simulations
do not have enough systems that resemble the LG features.
This is also standarized method to compare different simulations and
is also convenient for reproducibility and comparison 
We demonstrate the method using the Illustris-1 and ELVIS simulations.
We find significant differences between the two simulations. 
In the case of Illustris-1 $0.05\%$ to $0.25\%$ of the pairs with isolation
and dynamical characteristics similar to the LG are expected to have
satellite distributions with the same degree of deviation from
sphericity.
In the ELVIS simulations this fractions drops dramatically to
$0.001\%$ of the pairs.
This makes the LG a $3\sigma$ outlier in LCDM simulations, with all
the weight lying on the extremly planar MW satellite distribution.

@@

\begin{figure}
\centering
\includegraphics[width=0.47\textwidth]{expected_numbers.pdf}
\includegraphics[width=0.47\textwidth]{expected_numbers_flip.pdf}
\caption{Probability distribution for the expected number of pairs
  showing the same degree of atipicality as the Local Group if drawn
  from a sample of 2000 isolated pairs. 
  The distributions correspond to results derived from Illustris-1
  and ELVIS data.
  On average between $0.05\%$ and $0.25\%$ of the isolated pairs should present
  satellite distributions as atypical as the Local Group.
\label{fig:expected_number}}
\end{figure}
